---
title: "HW 3: Visualization, Ggplot2 & Maps"
subtitle: "Data 180, Professor Bilen"
author: 
  name: "XXXX"
  email: "XXXX@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** October 11 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: This time we will practice visualization using ggplot2, and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(ggplot2) # keep in mind you have to install the package if it's not installed in your machine
library(MASS)
data(Cars93)
head(Cars93)
```

# Ggplot2

## Task 1
Create a simple scatterplot using ggplot2 where x axis shows `EngineSize`, y axis shows `MPG.city`.

* Use blue color for your dots.
* Use size = 2 for the dot sizes.
* Use xlim and ylim to make sure your x and y axis ranges make sense.
* Add a title, subtitle, x axis label, y axis label, and a caption using title, subtitle, caption options inside `+ labs()`.

 
```{r}
library(ggplot2)
ggplot(Cars93, aes(x=Weight,y=MPG.city)) + 
                geom_point(col='blue') + labs(title = "MPG City by Weight",
              subtitle = "Each point corresponds to a car",
              caption = "Data source: Cars93")
```

## Task 2
Recreate the scatterplot in Task 1. Add the following functionality:

* Use a different color for each car `Type`.
```{r}
library(ggplot2)
ggplot(Cars93, aes(x=Weight,y=MPG.city)) + 
                geom_point(aes(color=Type))
```

## Task 3
Recreate the scatterplot in Task 2. Add the following functionality:

* Fit and overlay a smooth curve.
* Make the curve line red, and dashed. (Hint linetype is the option you can use to set line style as 'dashed')

```{r}
library(ggplot2)
ggplot(Cars93, aes(x=Weight,y=MPG.city)) + 
                geom_point(aes(color=Type)) +
                geom_smooth(color='red',linetype = "dashed") 
```

## Task 4
Create a barplot showing count of cars by `AirBags` using ggplot2.

* `geom_bar()` is the function that creates a barplot.
* Use a different color for each bar.
* Add a title, subtitle, x axis label, y axis label, and a caption using title, subtitle, caption options inside `+ labs()`.

```{r}
ggplot(Cars93,aes(x=AirBags)) +
      geom_bar(aes(fill=AirBags)) + labs(title = "Number of cars by AirBag",
              subtitle = "Each bar shows the count of cars",
              caption = "Data source: Cars93")
```

## Task 5
Recreate the barplot in Task 4 for different car `Origin`, such that you create one barplot for each `Origin` category and display all on the same graph.

* You can have `Origin` as the "x" variable by setting `x=Origin.`
* Use `fill=AirBags` to get a different color (and bar) by `AirBags`.
* in `geom_bar(), use `position='dodge'` to have the barplots side by side.

```{r}
ggplot(Cars93,aes(x=Origin)) +
      geom_bar(aes(fill=AirBags),position='dodge')
```

# Maps
Load Australia state map data via running:
```{r}
# install the packages below by uncommenting the lines. You need to do this only once.
#install.packages('maps')
#install.packages('ozmaps')
#install.packages('sf')
library(maps)
library(ozmaps)
library(sf)
au_states <- ozmaps::ozmap_states

# Display the data
head(au_states)
```

## Task 1
Create a map of Australia using ggplot.
```{r}
ggplot(au_states) + 
  geom_sf()
```

## Task 2
Modify your map such that each state is displayed with a different color. (Hint: use the fill option as an aes setting with variable `NAME`.)
```{r}
ggplot(au_states) + 
  geom_sf(aes(fill = NAME))
```
## Task 3
Read in the capital city dataframe by running:
```{r}
au_capitals <- tibble::tribble( 
  ~city,           ~lat,     ~lon,
  "Sydney",    -33.8688, 151.2093,  
  "Melbourne", -37.8136, 144.9631, 
  "Brisbane",  -27.4698, 153.0251, 
  "Adelaide",  -34.9285, 138.6007, 
  "Perth",     -31.9505, 115.8605, 
  "Hobart",    -42.8821, 147.3272, 
  "Canberra",  -35.2809, 149.1300, 
  "Darwin",    -12.4634, 130.8456, 
)
```

Display the capitals as points on your map from Task 1. (Hint: combine geom_sf() with geom_point())

```{r}
ggplot() + 
  geom_sf(data = au_states, colour = "black", fill = NA) + 
  geom_point(data = au_capitals, aes(x = lon, y = lat), col = "red", size=2)
```


## Task 4
Let's take a look at different projection output. Change your map's projection to epsg 3112 by adding `+ coord_sf()`. Recall the option that goes inside is `st_crs(4_digit_code)`
```{r}
ggplot(au_states) + geom_sf()
ggplot(au_states) + geom_sf() + coord_sf(crs = st_crs(3112))
```

## Task 5
Lastly, let's display the same map using mapview()

```{r}
#install.packages('mapview')
library(mapview)
mapview(au_states)
# you can use the mapview options to turn off legends, hover etc: legend = FALSE, label = FALSE, popup = FALSE
```

## Task 6
Let's display the capitals on the map like before. First convert the `au_capitals` dataframe to spatial dataframe object by running:
```{r}
capitals_geo <- st_as_sf(au_capitals, coords = c("lon", "lat"))
```

Now add the new spatial capitals object via + (that is, you can use the following syntax to add a layer: `mapview() + mapview()`)

```{r}
mapview(au_states) + mapview(capitals_geo,col.regions='red',legend=F)
```



You are done! ðŸ Don't forget to commit and push your .Rmd file to your Github repository before the due date.


