---
title: "HW 1"
subtitle: "Data 180, Professor Bilen"
author: 
  name: "Answer key"
  email: "XXXX@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document
---
Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 19 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

## Preliminary Task
Go to [datacamp.com](https://datacamp.com) and complete the tutorial "Introduction to R".


## Task 1
a) Store the data below in a data frame object named `my_df`.


|Person | Preference|
|:------|:----------|
|Alice | Twitter |
| Bob | Facebook |
| Charlie | Twitter |
| Denise | Instagram |
| Erin | Facebook |
| Frank | Twitter |
| Gisele | Twitter |
| Harry | Instagram |

```{r}
my_df <- data.frame(Person=c('Alice','Bob','Charlie','Denise','Erin','Frank','Gisele','Harry'),Preference=c('Twitter','Facebook','Twitter','Instagram','Facebook','Twitter','Twitter','Instagram'))
my_df
```

b) What are the dimensions of `my_df`?
```{r}
dim(my_df)
```

`my_df` has 8 rows, 2 columns.

c) Write R code that counts the total number of users.
```{r}
# Given each row corresponds to a unique unit,
nrow(my_df)
```

d) Write R code that counts the total number of users who prefer Twitter.
```{r}
# This requires subsetting. We need a way to figure out which rows have Twitter preference:
my_df$Preference == 'Twitter' # This returns a boolean vector, i.e. TRUE when preference is Twitter. We can use this vector for subsetting:
my_df[my_df$Preference=='Twitter',] # Notice we want to only subset rows, so don't forget to put a comma indicating that we want to be returned all the columns.
```

e) Write R code that returns what Gisele prefers.
```{r}
# It's fine if you did the following:
my_df[7,]
# A more generalizable approach would be:
my_df[my_df$Person=='Gisele',]
```

## Task 2
a) Create a vector named `my_vec` that includes all integers from 1 to 10 (inclusive).
```{r}
my_vec = 1:10
```

b) Replace the second element in `my_vec` with the number `99`.
```{r}
my_vec[2] = 99
```

c) Write a function called `my_func` that prints `This element is greater than 5!` if an element in `myvec` is greater than 5, and `This element is NOT greater than 5!` otherwise.
```{r}
# Write a function that takes an integer as an input, returns the correct prompt depending on the number being < or > 5.
my_func = function(num){
  if(num>5){
    print('This element is greater than 5!')
  } else{
    print('This element is NOT greater than 5!')
  }
}
# Test the function
my_func(my_vec[1])
my_func(my_vec[2])
my_func(my_vec[3])
```

## Task 3
a. Write a function named `get_fibonacci` that returns the n'th value in the [fibonacci sequence](https://www.cuemath.com/numbers/fibonacci-sequence). For example, `get_fibonacci(5)` should return 5; `get_fibonacci(7)` should return 13 and so on. ***Hint:** It is possible to call the same function from inside it. E.g., you can call `get_fibonacci()` from inside `get_fibonacci()`. Like the movie inception!*

```{r}
get_fibonacci = function(num){
  
  if(num==0){
    return(0)
  } else if (num==1){
    return(1)
  } else {
    return(get_fibonacci(num-1) + get_fibonacci(num-2))
  }
}
get_fibonacci(5)
get_fibonacci(7)
```

b. Using your function in part (a), find and return the 19th value in the fibonacci sequence.
```{r}
get_fibonacci(19)
```
<br>
You are done! 🏁 Don't forget to `commit` and `push` your .Rmd file to your Github clone repository before the due date.
<br>

