---
title: "Lecture 4 cont: ggplot2"
output: 
  html_document
---


```{r echo=FALSE}
# Custom options for knitting
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "#",
  fig.height = 5,
  fig.width = 5,
  fig.align = "center",
  cache = FALSE
) 
```

# ggplot2 R package for Data Visualization  


* ggplot is a function in the [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html) package and is based on **The Grammar of Graphics** by Leland Wilkinson, and the [lattice](https://cran.r-project.org/web/packages/lattice/index.html) package
* ggplot is designed to work in a  *layered fashion*, starting with a layer showing the raw data then adding layers of annotation and statistical summaries 

## ggplot() function
Let's look at an example **diamonds** data that comes with the [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html) package.  But first, let's load the the package to our session using the **library()** function.  If you have not yet installed the ggplot2 package, you should do this first (you only have to do this once). Recall the magic command is: install.packages("")


```{r echo=FALSE, warnings=FALSE}
#install.packages("ggplot2") # only need to install once
library(ggplot2)
```

Some information about the **diamonds** dataset :

* ~54,000 round diamonds from http://www.diamondse.info/
* Variables:
	+ carat, colour, clarity, cut
	+ total depth, table, depth, width, height
	+ price 

* A question of interest: What is the relationship between *carat* and *price*, and how does it depend on other factors?


## Data on diamonds 

Load the **diamonds** data set, get the dimensions, and look at the first few lines

```{r, warnings=FALSE}
data(diamonds, package="ggplot2")
dim(diamonds)
head(diamonds)
```


## Refresher: Default R plot of price versus carat

Recall using the *default settings* with the **plot()** function
```{r, fig.height=5, fig.width=5, fig.cap="Here is a caption."}
plot(diamonds$carat,diamonds$price) # x-variable first in this notation
# or
plot(price~carat, data=diamonds) # an alternative way: this is y against x
```

Note: If you want to change the figure size and add a figure caption in Rmd, you can specify the fig.height, fig.width, fig.cap options inside curly brackets at the beginning of your R code chunk. See the full set of options [here](https://rpubs.com/Lingling912/870659) Bonus: If you don't want to set the size each time you generate a plot, you can insert the options, e.g., fig_width fig_height in your YAML chunk at the beginning of your Rmd file.

## Default ggplot price versus diamonds

Here comes ggplot! Using the *default settings* in **ggplot()**

```{r}
theme_set(theme_bw())
library(ggplot2)
ggplot(diamonds, aes(x=carat,y=price)) + 
                          geom_point() + 
                          labs(y = 'price $', x = 'carat value')
```

## At first glance

* The *default* option in **ggplot** looks nicer!
* **ggplot**'s syntax looks weird, especially if you're not very familiar with **lattice**
* Note: It is possible to manipulate the **plot()** options to get a plot similar to ggplot (and maybe even a better one), but that would require a lot of extra coding!!
* The ggplot *syntax* also makes plotting more  **structured** and  easier to update 


## ggplot syntax

ggplot(diamonds, aes(x=carat,y=price)) + geom_point()

 
*The first component (before the "+") calls the ggplot function, and the data with x-y varibles.
* The second component (after the first "+") tells ggplot what type of plot you want, e.g., geom_bar/geom_hist/geom_boxplot
* Possible to add other lines for more customization on the plot, e.g., title, label for the axes, etc.
** https://rkabacoff.github.io/datavis/  a whole book on ggplot2 customization!!


##  geom_boxplot
Very easy to switch to a boxplot. We can use  **geom_boxplot** to create boxplots when one variable is continuous and the other is a factor. 

```{r,}
ggplot(diamonds, aes(x=cut,y=price)) + geom_boxplot()
```


## Changing the aesthetics

You can control the aesthetics of each layer, e.g.  *colour*, *size*, *shape*, *alpha* (opacity) etc.

```{r}
ggplot(diamonds, aes(carat, price)) + geom_point(col = "blue")
```


## A few more examples
Changing the alpha level

```{r}
ggplot(diamonds, aes(x=carat,y=price)) + geom_point(alpha = 0.2)
```


## A few more examples

Changing the point size

```{r}
ggplot(diamonds, aes(x=carat,y=price)) + geom_point(size = 0.2)

```


## A few more examples

Changing the shape and the point size
```{r}
ggplot(diamonds, aes(x=carat,y=price)) + geom_point(shape = 2,size=0.4)
```

 
## Combining layers
The real power of **ggplot** is its ability to combine layers
```{r}
ggplot(diamonds, aes(x=carat,y=price)) + geom_point(size = 0.2) +
  geom_smooth()
```


## Transformations

In this case (and many other situations) a log transformation may allow for the relationships between variables to be clearer. Can use **coord_trans()**

```{r}
ggplot(diamonds, aes(carat, price)) + geom_point(size = 0.5) +
coord_trans(x = "log10", y = "log10")
```

## Adding information for a third variable
We can color by a *factor* variable (not that it's useful for this example!)

```{r}
ggplot(diamonds, aes(carat, price, colour=color)) + geom_point() + 
    coord_trans(x = "log10", y = "log10")
```


Can also color by a  *continuous* variable (not really useful for this example too, but here it is so you are familiar with the syntax:)
```{r}
ggplot(diamonds, aes(carat, price, colour=depth)) + geom_point() + 
    coord_trans(x = "log10", y = "log10")
```

## Conditional plots:

In some cases, it may be more useful to get separate plots for each category of the third variable, to understand conditional relationships 

```{r}
ggplot(diamonds, aes(carat, price)) + geom_point() +
  facet_wrap(~color, ncol=4)
```


Alternatively, you can use **facet_grid**, which also allows more than 1 conditioning variable (tables of plots)
```{r, fig.height=3, fig.width=8}
ggplot(diamonds, aes(carat, price)) + geom_point() +
 facet_grid(~color, labeller=label_both)
```


## A final note about syntax
There are actually many ways to get the same plot! The following commands will produce the same plot:

* ggplot(diamonds, aes(price, carat)) + geom_point()
* ggplot() + geom_point(aes(price, carat), diamonds)
* ggplot(diamonds) + geom_point(aes(price, carat))
* ggplot(diamonds, aes(price)) + geom_point(aes(y = carat))
* ggplot(diamonds, aes(y=carat)) + geom_point(aes(x = price))

## A final summary of syntax:

```{r}
ggplot(diamonds) + geom_point(aes(price, carat))
```


## Histograms
Let's make a histogram.
```{r}
ggplot(diamonds, aes(depth)) + geom_histogram()
```

Notice the difference in the **aes** call; **boxplot** is really designed for multiple categories!

## Histograms
Tthe *default options* in **histogram** may not be sensible, and you often need to adjust the **binwidth** and **xlim**

```{r}
ggplot(diamonds, aes(depth)) + geom_histogram(binwidth=0.2) + xlim(56,67)
```

## Boxplots with multiple categories
A better use of **boxplot** is when we want to compare distributions of a quantitative variable across categories of a factor variable, as previously discussed

```{r}
ggplot(diamonds, aes(cut, depth)) + geom_boxplot()
```



## Histograms with multiple categories
We can also get multiple histograms, though we need to either display them separately (less useful when comparing)
```{r}
ggplot(diamonds, aes(depth)) + geom_histogram(binwidth = 0.2) + 
    facet_wrap(~cut) + xlim(56, 67)
```

## Overlaying Histograms
Or, you can overlay the historgrams 
```{r}
ggplot(diamonds, aes(depth, fill=cut)) + 
    geom_histogram(binwidth=0.2,colour="grey50",alpha=.4,position="identity") + xlim(56,67)
```

## Cheat sheet
We are covering only a few of the many plot types that can be greated with the **ggplot2** package.

For a more comprehensive view of ggplot2, take a look at the ggplot2 [Cheat sheet](https://r4ds.had.co.nz/data-visualisation.html)


## Maps
### load United States state map data
```{r}
#install.packages('maps') # you only need to do this once. maps package includes various maps that we can use.
#install.packages('sf') # you only need to do this once
library(maps)     # Provides latitude and longitude data for various maps
library(sf)
# read the state population data
MainStates <- map_data("state")

#plot all states with ggplot2, using black borders and light blue fill
ggplot() + 
  geom_polygon( data=MainStates, aes(x=long, y=lat, group=group),
                color="black", fill="lightblue" ) +
                coord_sf(crs = st_crs(4326)) # projection

```


```{r}
library(mapview)
library(sf)
states <- st_read('/Users/bilene/Downloads/us-states/us-states.shp')
# you can select the map styles (alternatively: map.types = "Esri.WorldImagery" add option inside mapview)
# mapviewOptions(basemaps = c("OpenStreetMap","Esri.WorldImagery","OpenTopoMap"))
mapview(states,zcol='population', col.regions = blues9)
```

```{r}
map1 <- mapview(states[states$NAME=='Pennsylvania',], col.regions = blues9, layer.name = "PA") 
map2 <- mapview(states[states$NAME=='South Carolina',], col.regions = blues9, layer.name = "SC")
# use the mapview options to turn off legends, hover etc: legend = FALSE, label = FALSE, popup = FALSE or more options such as label=paste('NAME:',states$NAME,'<br> avg. age:',states$age) are possible.
#install.packages('leafsync')
library(leafsync)
sync(map1, map2)
#or 
#mapviewOptions(basemaps = c("OpenStreetMap"))
#map1 | map2
```
